---
title: "Initial Tony Analysis"
author: "Maximilian Wisnefski"
date: '2023-10-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(magrittr)
library(car)
library(ggpubr)
```


## Data Wrangling

Before we can do any analysis, there is some filtering and wrangling that need to be done on this data for it to work for the intended analysis

```{r}
tony_data <- read_csv("data/tony_data.csv", show_col_types = FALSE)
weekly_data <- read_csv("data/weekly_data.csv", show_col_types = FALSE)
```


First, we need to remove any rows with missing data (only applies to 12 rows of data)

```{r}
#removing rows that have missing data
weekly_data = weekly_data[weekly_data$capacity > 0, ] 
weekly_data = weekly_data[weekly_data$gross > 0, ] 

#the number of unique productions that we have gross/capacity data on from 1979 - 2023
length(unique(weekly_data$production))
```


```{r}
#the values for date are all strings. We need to change them to vars of class date
weekly_data %<>%
  mutate(date = as.Date(date, format= "%B %d, %Y"))
```

```{r}
#the first merged_data is an inner join, the second is an outer
merged_data <- merge(weekly_data, tony_data)
merged_data_all <- merge(weekly_data, tony_data, all = TRUE)
```


If a production closed prior to the nominations announcement in May, then there was no way for that production to be affected positively or negatively by the awards. Therefore, it ought to be removed from the data. 

```{r}
#filter out productions that did not fall within time range of Tony nom announcements and ceremony (roughly May 1 - June 15 for all years)
prods_in_range_df <- merged_data_all %>%
  filter(date >= as.Date(paste(year(date), 05, 01, sep = "-")),
         date <= as.Date(paste(year(date), 06, 15, sep = "-")))

#this is a list of productions that were open within the period of time during which they could be affected by Tony wins and loses
prods_in_range <- unique(prods_in_range_df$production)
#prods_in_range
```


```{r}
#select all rows where production is in list of productions of interest
all_relevant_data <- merged_data_all %>% filter(production %in% prods_in_range)
length(unique(all_relevant_data$production))
```

As you can see, the number of productions in our data decreased by almost half. Many productions do not actually make it to their season's Tony Awards. This can happen because the production is a limited run (a production that is only supposed to run for a set number of weeks, regardless of financial success), or because the production is a flop (it closes after a very short time due to lack of commercial/critical success). Note: this number will actually go down to 837 soon because of some additional filtering.


Now we need to fix the 'year' column for rows where productions received 0 nominations. The 'year' column is supposed to represent the year in which a production was eligible for the Tony Awards. However, when web scraping, I realized that productions that received 0 noms did not have this info explicitly available on the database, so instead I scraped the date that the production opened. To correct this issue, we need to change this date to the correct year of eligibility. A Tony season runs from early May of one year to late April of the next. Therefore, if a production opened on, say, January 12, 2022, it was eligible for the 2022 Tony Awards, but if a production opened on August 15, 2022, it was eligible for the 2023 Tony Awards. 

```{r}
#code to fix 'year' column issue mentioned above
for (i in 1:length(all_relevant_data$year)){
  year <- all_relevant_data$year[i]
  if (nchar(year) > 4){
    d <- as.Date(year, format = "%B %d, %Y")
    
    #if date is later than start of May
    if (d >= as.Date(paste(year(d), 05, 01, sep = "-"))){
      #it is eligible for Tonys during next calender year
      actual_year <- as.integer(format(as.Date(d, format="%d/%m/%Y"),"%Y")) + 1
      all_relevant_data$year[i] <- actual_year
    }
    
    #if date is earlier than end of April
    else{
      #it is eligible for Tonys in current calender year
      actual_year <- as.integer(format(as.Date(d, format="%d/%m/%Y"),"%Y"))
      all_relevant_data$year[i] <- actual_year
      #year = year
    }
  }
  #converting date string into an int
  else{
    all_relevant_data$year[i] <- as.integer(all_relevant_data$year[i])
  }
}
```


```{r}
#get rid of productions that are eligible for the 2024 Tonys as these have not happened yet (only applies to 2 productions in the data)
#get rid of productions that were eligible for the 2020 (actually 2021) Tonys as the Tonys were odd that year due to Covid (see final paper for more detailed explanation)
#get rid of productions that were eligible for pre-1980 Tonys as the data are too sparse
all_relevant_data <- all_relevant_data[all_relevant_data$year != 2024, ]
all_relevant_data <- all_relevant_data[all_relevant_data$year != 2020, ]
all_relevant_data <- all_relevant_data[all_relevant_data$year >= 1980, ]
length(unique(all_relevant_data$production))
```


```{r}
#only includes productions that had chance to be affected by Tonys
prod_longevity <- all_relevant_data %>% group_by(production)  %>%
                            summarise(weeks = n(), 
                                      nominations = mean(nominations), 
                                      wins = mean(wins),
                                      year = max(year), 
                                      type = min(type), 
                                      .groups = 'drop')

#only need this for summary statistics (want to include all productions, even flops and limited engagements)
prod_longevity_all_prods <- merged_data_all %>% group_by(production)  %>%
                            summarise(weeks = n(), 
                                      nominations = mean(nominations), 
                                      wins = mean(wins),
                                      year = max(year), 
                                      type = min(type), 
                                      .groups = 'drop')

```


## Summary Statistics

We need to calculate and visualize some basic summary statistics to better understand the data

```{r}
#average number of weeks that a Broadway production runs (play, musical, special)
avg_run_len <- mean(prod_longevity_all_prods$weeks)
avg_run_len
```

```{r}
#median longevity
median_run_len <- median(prod_longevity_all_prods$weeks)
median_run_len
```


```{r}
ggplot(prod_longevity, aes(x = weeks)) +
  geom_histogram(fill = "skyblue", color = "black") +
  geom_vline(xintercept = 14,lwd=1,colour="maroon", linetype = 2) +
  xlim(0,450) +
  labs(x = "runtime (weeks)") +
  ggtitle("Distribution of Runtimes for All Relevant Productions") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
#number of productions that are relevant to analysis for each year (i.e. Broadway season)
prods_per_year <- prod_longevity %>% group_by(year) %>%
                        summarise(num_prods = n())

mean(prods_per_year$num_prods)
```


```{r}
#average capacity for each month
#these results are slightly different from the early results now that we've scraped more data

#creating column that just has month of year
merged_data_all$month <- format(as.Date(merged_data_all$date, format="%d/%m/%Y"),"%m")
monthly_capacity <- merged_data_all %>% 
    #group_by(month = lubridate::floor_date(date, 'month')) %>%
    group_by(month) %>%
    summarize(avg_capacity = format(mean(capacity), scientific = FALSE))
monthly_capacity
```


## Initial Analysis

```{r}
#remove specials since they are limited engagements (their longevity cannot be affected by Tonys)
prod_longevity <- prod_longevity[prod_longevity$type != "special", ]

noms_longev_plot <- ggplot(prod_longevity, aes(x = wins, y = weeks)) +
  geom_jitter(color = 'maroon') +
  #not sure why line of best fit won't show up
  geom_smooth(method = "lm", color = 'skyblue') +
  ylim(0,450) +
  labs(x = "Tony wins", y = "weeks running") +
  ggtitle("Effect of Tony Wins on Production Longevity") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))

wins_longev_plot <- ggplot(prod_longevity, aes(x = nominations, y = weeks)) +
  geom_jitter(color = 'maroon') +
  #not sure why line of best fit won't show up
  geom_smooth(method = "lm", color = 'skyblue') +
  ylim(0,450) +
  labs(x = "Tony nominations", y = "weeks running") +
  ggtitle("Effect of Tony Nominations on Production Longevity") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(wins_longev_plot, noms_longev_plot, ncol = 1, nrow = 2, labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/noms_wins_longevity.png", bg = "transparent", height = 10, width = 10)
```


```{r}
longevity_lm <- lm(weeks ~ nominations + wins, data = prod_longevity)
summary(longevity_lm)
```


```{r}
#calculating confidence interval
confint(longevity_lm, level=0.95)
```


Note that the output of the regression model is distinct from the lines of best fit in the two scatterplots above. Both scatterplots are simple linear regression (one predictor), whereas the model is multiple linear regression.

```{r}
#this indicates that there is a small correlation between nominations and wins (obviously), but it is not large enough to be a serious issue of multicollinearity
vif(longevity_lm)
```


### Checking assumptions of the Linear Regression model


```{r}
plot(longevity_lm)
```


```{r}
#does not meet homoscedasticity assumption
ggplot(longevity_lm, aes(x = longevity_lm$fitted.values, y = longevity_lm$residuals)) +
  geom_point(color = "maroon") +
  geom_hline(yintercept = 0, color = "skyblue") +
  labs(x = "fitted values", y = "residuals") +
  ggtitle("Residuals vs Fitted\n(for intial regression model)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) 
```



```{r}
#fairly normally distributed, but not at the ends
QQ_plot <- ggplot(longevity_lm, aes(sample = longevity_lm$residuals)) +
  stat_qq(color = "maroon") +
  stat_qq_line(color = "skyblue") + 
  labs(x = "theoretical quantiles", y = "sample quantiles") +
  ggtitle("Normal Q-Q Plot\n(for intial regression model)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

resid_hist <- ggplot(longevity_lm, aes(x = longevity_lm$resid)) +
  geom_histogram(color = "black", fill = "skyblue") +
  xlim(-500,500) +
  labs(x = "residual value", y = "frequency") +
  ggtitle("Distribution of Residuals\n(for intial regression model)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(QQ_plot, resid_hist, ncol = 1, nrow = 2, labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/QQ+hist.png", bg = "transparent", height = 10, width = 8)
```


## Time Series for Two Individual Productions

Here we look at two scenarios from the 2006 - 2007 season: one in which a production (Spring Awakening) received 11 nominations and won 8, and another in which a production (The Pirate Queen) received 1 nomination and won 0. The goal is to anecdotally illustrate how the awards can impact a production's financial success in the weeks following the nominations announcement, and the weeks following the ceremony.

```{r}
#make dataframes with weekly data for only those productions
spring_awake <- weekly_data[weekly_data$production == 'spring-awakening-448811',] 
pirate_queen <- weekly_data[weekly_data$production == 'the-pirate-queen-426183',] 
```

```{r}
spring_awake_plot <- ggplot(spring_awake, aes(x = date, y = gross))+
  geom_line(color = 'maroon') +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%B",
                 limits = as.Date(c('2006-11-19','2007-08-05'))) +
  labs(y = "weekly gross") +
  ggtitle("Change in Weekly Gross for Spring Awakening\n(2006 - 2007 season)") +
  theme_bw() +
  geom_vline(xintercept =  as.numeric(as.Date("2007-05-14")), lwd = 1, color = 'skyblue') +
  geom_vline(xintercept =  as.numeric(as.Date("2007-06-10")), lwd = 1, color = 'skyblue') +
  theme(plot.title = element_text(hjust = 0.5))


pirate_queen_plot <- ggplot(pirate_queen, aes(x = date, y = gross))+
  geom_line(color = 'maroon') +
  ylim(100000, 900000) +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%B",
                 limits = as.Date(c('2007-03-11','2007-08-05'))) +
  labs(y = "weekly gross") +
  ggtitle("Change in Weekly Gross for The Pirate Queen\n(2006 - 2007 season)") +
  theme_bw() +
  geom_vline(xintercept =  as.numeric(as.Date("2007-05-14")), lwd = 1, color = 'skyblue') +
  geom_vline(xintercept =  as.numeric(as.Date("2007-06-10")), lwd = 1, color = 'skyblue') +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(spring_awake_plot, pirate_queen_plot, ncol = 1, nrow = 2, labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/two_shows_ex.png", bg = "transparent", height = 10, width = 8)
```


```{r}
#save data needed for next steps
#write.csv(all_relevant_data, "data/combined_filtered.csv", row.names = FALSE)
#write.csv(prod_longevity, "data/production_longevity.csv", row.names = FALSE)
```



