---
title: "Initial Tony Analysis"
author: "Maximilian Wisnefski"
date: '2023-10-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(magrittr)


#data(mtcars)
#cars <- as.data.frame(mtcars)
```


TO DO: Add comments and markdown text!

## Data Wrangling

```{r}
tony_data <- read_csv("data/tony_data.csv", show_col_types = FALSE)
weekly_data <- read_csv("data/weekly_data.csv", show_col_types = FALSE)
```

```{r}
#removing rows I that have missing data
weekly_data = weekly_data[weekly_data$capacity > 0, ] 
weekly_data = weekly_data[weekly_data$gross > 0, ] 

#the number of unique productions that I have gross/capacity data on from 1979 - 2023
length(unique(weekly_data$production))
```


The above number is higher than the number of rows in tony_data. I am almost positive that this is the case because my web scraping code for tony_data ignored productions that did not receive any nominations. I am basing this off of the fact that "Bad Cinderella" (which is in weekly_data) did not receive any nominations and does not appear in tony_data. This applies to a few other productions as well

```{r}
#the values for date are all strings. I need to change them vars of class date
weekly_data %<>%
  mutate(date = as.Date(date, format= "%B %d, %Y"))
```

```{r}
#the first merged_data is an inner join, the second is an outer
merged_data <- merge(weekly_data, tony_data)
merged_data_all <- merge(weekly_data, tony_data, all = TRUE)
```


```{r}
#filter out productions that did not fall within time range of Tony nom announcements and ceremony (roughly May 1 - June 15 for all years)
prods_in_range_df <- merged_data_all %>%
  filter(date >= as.Date(paste(year(date), 05, 01, sep = "-")),
         date <= as.Date(paste(year(date), 06, 15, sep = "-")))

#this is a list of productions that were open within the period of time during which they could be affected by Tony wins and loses
prods_in_range <- unique(prods_in_range_df$production)
#prods_in_range
```


```{r}
#select all rows where production is in list of productions of interest
all_relevant_data <- merged_data_all %>% filter(production %in% prods_in_range)
length(unique(all_relevant_data$production))
```

As you can see, the number of productions in my data decreased by almost half. Many productions do not actually make it to their season's Tony Awards. This can happen because the production is a limited run (a production that is only supposed to run for a set number of weeks, regardless of financial success), or because the production is a flop (it closes after a very short time due to lack of commercial/critical success). Note: this number will actually go down to 837 soon because of some additional filtering.


Now I need to fix the 'year' column for rows where productions received 0 nominations. The 'year' column is supposed to represent the year in which a production was eligible for the Tony Awards. However, when web scraping, I realized that productions that received 0 noms did not have this info explicitly available on the database, so instead I scraped the date that the production opened. To correct this issue, I need to change this date to the correct year of eligibility. A Tony season runs from early May of one year to late April of the next. Therefore, if a production opened on January 12, 2022, it was eligible for the 2022 Tony Awards, but if a production opened on August 15, 2022, it was eligible for the 2023 Tony Awards. 

```{r}
#code to fix 'year' column issue mentioned above
for (i in 1:length(all_relevant_data$year)){
  year <- all_relevant_data$year[i]
  if (nchar(year) > 4){
    d <- as.Date(year, format = "%B %d, %Y")
    
    #if date is later than start of May
    if (d >= as.Date(paste(year(d), 05, 01, sep = "-"))){
      #it is eligible for Tonys during next calender year
      actual_year <- as.integer(format(as.Date(d, format="%d/%m/%Y"),"%Y")) + 1
      all_relevant_data$year[i] <- actual_year
    }
    
    #if date is earlier than end of April
    else{
      #it is eligible for Tonys in current calender year
      actual_year <- as.integer(format(as.Date(d, format="%d/%m/%Y"),"%Y"))
      all_relevant_data$year[i] <- actual_year
      #year = year
    }
  }
  #just converting date string into an int
  else{
    all_relevant_data$year[i] <- as.integer(all_relevant_data$year[i])
  }
}
```


```{r}
#get rid of productions that are eligible for the 2024 Tonys as these have not happened yet (only applies to 2 productions in the data)
#get rid of productions that are eligible for the 2020 (actually 2021) Tonys as the Tonys were odd that year due to Covid (will discuss this more in final paper)
#get rid of productions that were eligible for pre-1980 Tonys as the data are too sparse
all_relevant_data <- all_relevant_data[all_relevant_data$year != 2024, ]
all_relevant_data <- all_relevant_data[all_relevant_data$year != 2020, ]
all_relevant_data <- all_relevant_data[all_relevant_data$year >= 1980, ]
length(unique(all_relevant_data$production))
```


```{r}
#only includes productions that had chance to be affected by Tonys
prod_longevity <- all_relevant_data %>% group_by(production)  %>%
                            summarise(weeks = n(), 
                                      nominations = mean(nominations), 
                                      wins = mean(wins),
                                      year = max(year), 
                                      type = min(type), 
                                      .groups = 'drop')

#only need this for summary statistics (want to include all productions, even flops and limited engagements)
prod_longevity_all_prods <- merged_data_all %>% group_by(production)  %>%
                            summarise(weeks = n(), 
                                      nominations = mean(nominations), 
                                      wins = mean(wins),
                                      year = max(year), 
                                      type = min(type), 
                                      .groups = 'drop')

```


## Summary Statistics

```{r}
#average number of weeks that a Broadway production runs (play, musical, special)
#slightly different number than before because you added more prods in most recent scraping
avg_run_len <- mean(prod_longevity_all_prods$weeks)
avg_run_len
```

```{r}
#median longevity
#same number as before 
median_run_len <- median(prod_longevity_all_prods$weeks)
median_run_len
```


```{r}
#number of productions that are relevant to analysis for each year (i.e. Broadway season)
prods_per_year <- prod_longevity %>% group_by(year) %>%
                        summarise(num_prods = n())

mean(prods_per_year$num_prods)
```


```{r}
#average capacity for each month
#these results are slightly different now that you've added more productions

#creating column that just has month of year
merged_data_all$month <- format(as.Date(merged_data_all$date, format="%d/%m/%Y"),"%m")
monthly_capacity <- merged_data_all %>% 
    #group_by(month = lubridate::floor_date(date, 'month')) %>%
    group_by(month) %>%
    summarize(avg_capacity = format(mean(capacity), scientific = FALSE))
monthly_capacity
```


## Initial Analysis

```{r}
ggplot(prod_longevity, aes(x = wins, y = weeks)) +
  geom_jitter(color = 'maroon') +
  #not sure why line of best fit won't show up
  geom_smooth(method = "lm", color = 'skyblue') +
  ylim(0,450) +
  labs(x = "Tony wins", y = "weeks running") +
  ggtitle("Effect of Tony Wins on Production Longevity") +
  theme_bw() + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(prod_longevity, aes(x = nominations, y = weeks)) +
  geom_jitter(color = 'maroon') +
  #not sure why line of best fit won't show up
  geom_smooth(method = "lm", color = 'skyblue') +
  ylim(0,450) +
  labs(x = "Tony nominations", y = "weeks running") +
  ggtitle("Effect of Tony Nominations on Production Longevity") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
longevity_lm <- lm(weeks ~ nominations + wins, data = prod_longevity)
summary(longevity_lm)
```

### Checking assumptions of the Linear Regression model

```{r}
qqnorm(longevity_lm$resid, col = "maroon")
qqline(longevity_lm$resid, col = "skyblue", lwd = 2)
```


```{r}
hist(longevity_lm$resid, xlim = c(-500,500), breaks = 30, col = "skyblue")
```


```{r}
spring_awake <- weekly_data[weekly_data$production == 'spring-awakening-448811',] 
pirate_queen <- weekly_data[weekly_data$production == 'the-pirate-queen-426183',] 
```

```{r}
ggplot(spring_awake, aes(x = date, y = gross))+
  geom_line(color = 'maroon') +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%B",
                 limits = as.Date(c('2006-11-19','2007-08-05'))) +
  labs(y = "weekly gross") +
  ggtitle("Change in Weekly Gross for Spring Awakening\n(2006 - 2007 season)") +
  theme_bw() +
  geom_vline(xintercept =  as.numeric(as.Date("2007-05-14")), lwd = 1, color = 'skyblue') +
  geom_vline(xintercept =  as.numeric(as.Date("2007-06-10")), lwd = 1, color = 'skyblue') +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
ggplot(pirate_queen, aes(x = date, y = gross))+
  geom_line(color = 'maroon') +
  ylim(100000, 900000) +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%B",
                 limits = as.Date(c('2007-03-11','2007-08-05'))) +
  labs(y = "weekly gross") +
  ggtitle("Change in Weekly Gross for The Pirate Queen\n(2006 - 2007 season)") +
  theme_bw() +
  geom_vline(xintercept =  as.numeric(as.Date("2007-05-14")), lwd = 1, color = 'skyblue') +
  geom_vline(xintercept =  as.numeric(as.Date("2007-06-10")), lwd = 1, color = 'skyblue') +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
#save data needed for next steps
#write.csv(all_relevant_data, "data/combined_filtered.csv", row.names = FALSE)
#write.csv(prod_longevity, "data/production_longevity.csv", row.names = FALSE)
```



