---
title: "Main Tony Analysis"
author: "Maximilian Wisnefski"
date: '2023-10-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(segmented)
```


```{r}
weekly_and_tony <- read_csv("data/combined_filtered.csv", show_col_types = FALSE)
production_longevity <- read_csv("data/production_longevity.csv", show_col_types = FALSE)
```


## Interrupted time series analysis

### formatting data 

```{r}
#adding a column for Julian date (might not be important, we'll see)
weekly_and_tony$julian_date <- julian(weekly_and_tony$date) #as.numeric(format(weekly_and_tony$date, "%j"))
```



To perform an interrupted time series analysis, you need two additional columns in the dataframe: One containing a binary variable that is 0 if the date is pre-nominations announcement and 1 if it is after, and another containing a binary variable that is 0 if the date is pre-ceremony date and 1 if it is after. 


```{r}
#TO DO (maybe): get exact dates of noms announcement and ceremony for every year instead of defaulting to May 1 and June 5

#adding this column for noms annoucement
weekly_and_tony$post_noms <- with(weekly_and_tony,
                                  ifelse(weekly_and_tony$date < as.Date(
                                    paste(weekly_and_tony$year, 05, 01, sep = "-")), 0, 1))

#adding this column for ceremony date
weekly_and_tony$post_awards <- with(weekly_and_tony,
                                  ifelse(weekly_and_tony$date < as.Date(
                                    paste(weekly_and_tony$year, 06, 05, sep = "-")), 0, 1))
```


```{r}
#add column for number of weeks a show has been open on any given week
#must order rows by date first 
weekly_and_tony <- weekly_and_tony[order(weekly_and_tony$date), ]

#add column 
weekly_and_tony$index <- ave(weekly_and_tony$production, FUN = seq_along) 
weekly_and_tony$weeks_open <- as.numeric(ave(weekly_and_tony$index, weekly_and_tony$production, FUN = seq_along))
#remove unnecessary index column
weekly_and_tony = subset(weekly_and_tony, select = -c(index))
```



Now I need to change the values of the "nominations" and "wins" columns to reflect the fact that prior to the date of the announcement, all productions had 0 nominations, and prior to the date of the ceremony, productions had 0 wins.

```{r}
#implementing this change
weekly_and_tony$nominations[weekly_and_tony$post_noms == 0] <- 0
weekly_and_tony$wins[weekly_and_tony$post_awards == 0] <- 0
```


```{r}
#want to analyze each year separately first
split_data_unclean <- split(weekly_and_tony, f = weekly_and_tony$year)
split_data <- list()

#splitting data changes column names for each year of data. To avoid having to deal with this later, we should change the column names back to what they originally were
cols_list <- colnames(weekly_and_tony)
for ( i in 1: length(split_data_unclean)){
  #change class to data frame for each year
  df <- as.data.frame(split_data_unclean[i])
  names(df) <- c(cols_list)
  #appending df to list where all dfs have the same column names
  split_data <- append(split_data, list(df))
}

#split_data
```


### analysis

Now we are ready to actually perform the interrupted time series analysis

```{r}
#testing this on a random year (2019)
test <- as.data.frame(split_data[40])
#names(test) <- c(cols_list)


#this graph seems too hectic to include in final report. Probably better to stick to more anecdotal things that you made in previous part, at least for visuals
ggplot(test, aes(x = date, y = capacity, color = production)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%b",
                 limits = as.Date(c('2018-05-01','2019-09-01'))) +
  #ggtitle("") +
  theme_bw() +
  theme(legend.position = "none")
```


```{r}
#Interrupted time series
#testing on just one production first
aint_proud <- test[test$production == 'aint-too-proud-520004',] 
aint_proud_tm <- lm(capacity ~ julian_date + post_noms + post_awards + weeks_open + nominations + wins, data = aint_proud)

#results indicate that for this show, the weeks it been open negatively affect its capacity (people want to see new things), time after noms led to a 4.8% increase in capacity, time after awards led to a 4.0% increase in capacity (not entirely accurate if you look at actual data for this show. Capacity did increase, but not by that much)
#did not deem number of Tony noms or wins relevant, probably because I am only looking at one show, so number of noms/wins almost never changes
summary(aint_proud_tm)
```

```{r}
#do the same thing as the above chunk but for every 2019-2020 production (still pre-covid)
test_tm <- lm(capacity ~ julian_date + post_noms + post_awards + weeks_open + nominations + wins, data = test)
summary(test_tm)
```


```{r}
test_lm <- lm(capacity ~ julian_date + nominations + wins + weeks_open, data = aint_proud)
segmented_test_lm <- segmented(test_lm, seg.Z = ~julian_date, 
                               psi = julian(as.Date("2019-05-01")))

#just for one random musical (Ain't too Proud)
summary(segmented_test_lm)
```



```{r}
#these results for segmented regression seem more accurate than they were before (i.e. when I did not include weeks_open. this var is very relevant to capacity)
test_lm <- lm(capacity ~ julian_date + nominations + wins + weeks_open, data = test)
segmented_test_lm <- segmented(test_lm, seg.Z = ~julian_date, 
                               psi = julian(as.Date("2019-05-01")))

#estimates breakpoint as being right after awards (June 15, 2019)
#this might be better option than time series?  
summary(segmented_test_lm)
```




```{r}
#how to convert Julian date back into regular date (the default origin is Jan 1, 1970)

#extracting breakpoint date from model 
x <- segmented_test_lm$indexU$julian_date
x2<- as.Date(x, origin = as.Date("1970-01-01"))
x2
```



```{r}
#TO DO: do longevity analysis for each year and make time series of results???
```



