---
title: "Main Tony Analysis"
author: "Maximilian Wisnefski"
date: '2023-10-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(segmented)
library(nlme)
```


```{r}
weekly_and_tony <- read_csv("data/combined_filtered.csv", show_col_types = FALSE)
production_longevity <- read_csv("data/production_longevity.csv", show_col_types = FALSE)
```


## Longevity analysis

```{r}
#production_longevity <- production_longevity[production_longevity$weeks <= 500, ]
production_longevity <- production_longevity[production_longevity$type != "special", ]
#want to analyze each year separately first
split_data_unclean <- split(production_longevity, f = production_longevity$year)
split_longev <- list()

#splitting data changes column names for each year of data. To avoid having to deal with this later, we should change the column names back to what they originally were
cols_list <- colnames(production_longevity)
for ( i in 1: length(split_data_unclean)){
  #change class to data frame for each year
  df <- as.data.frame(split_data_unclean[i])
  names(df) <- c(cols_list)
  #appending df to list where all dfs have the same column names
  split_longev <- append(split_longev, list(df))
}
```


```{r}
long_coeffs <- list()
long_signif <- list()
for (df in split_longev){
  longevity_lm <- lm(weeks ~ nominations + wins, data = df)
  coeffs <- c(longevity_lm$coefficients, df$year[1])
  long_coeffs <- append(long_coeffs, list(coeffs))

  p_vals <- summary(longevity_lm)$coefficients[,4] 
  r_sqr <- summary(longevity_lm)$adj.r.squared
  CI <- confint(longevity_lm, level=0.95)
  signif_list <- c(p_vals, r_sqr, CI, df$year[1])
  long_signif <- append(long_signif, list(signif_list))
  
}

#confint(longevity_lm,level=0.95)
```


```{r}
long_coeffs_df <- as.data.frame(do.call(rbind, long_coeffs))
colnames(long_coeffs_df) <- c("intercept_coeff", "noms_coeff", "wins_coeff", "year")

long_signif_df <- as.data.frame(do.call(rbind, long_signif))
colnames(long_signif_df) <- c("intercept_pval", "noms_pval", "wins_pval", 
                              "rsqr", "CI_1_int", "CI_1_noms", "CI_1_wins", 
                              "CI_2_int", "CI_2_noms", "CI_2_wins", "year")

long_lm_df <- merge(long_coeffs_df, long_signif_df)
```


```{r}
wins_effect_plot <- ggplot(long_lm_df, aes(x = wins_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_wins, xmax = CI_2_wins), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "wins effect size") +
  ggtitle("Effect Size of Wins on Longevity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

noms_effect_plot <- ggplot(long_lm_df, aes(x = noms_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_noms, xmax = CI_2_noms), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "nominations effect size") +
  ggtitle("Effect Size of Nominations on Longevity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(noms_effect_plot, wins_effect_plot, ncol = 1, nrow = 2,  labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/noms_wins_effect_longev.png", bg = "transparent", height = 10, width = 8)
```



It is important to note that the years that produced models with very large confidence intervals are all years that had an outlier production that stayed open for an unusually long time. 1988 has Phantom of the Opera, 1998 has The Lion King, 2004 has Wicked, and 2014 has Aladdin. 


```{r}
noms_effect_hist_longev <- ggplot(long_lm_df, aes(x = noms_coeff)) +
  geom_histogram(color = "black", fill = "skyblue") +
  geom_vline(xintercept = 0,lwd=.5,colour="maroon", linetype = 2) +
  ggtitle("Distribution of Yearly Regression Model Slopes for Nominations") +
  labs(x = "nominations effect size") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

wins_effect_hist_longev <- ggplot(long_lm_df, aes(x = wins_coeff)) +
  geom_histogram(color = "black", fill = "skyblue") +
  geom_vline(xintercept = 0,lwd=.5,colour="maroon", linetype = 2) +
  ggtitle("Distribution of Yearly Regression Model Slopes for Wins") +
  labs(x = "wins effect size") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(noms_effect_hist_longev, wins_effect_hist_longev, ncol = 1, nrow = 2,  labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/noms_wins_effect_hists.png", bg = "transparent", height = 10, width = 8)
```


```{r}
#many of the models only fit the data moderately well
mean(long_lm_df$rsqr)
median(long_lm_df$rsqr)
```


## Segmented regression and interrupted time series analysis

### formatting data 


```{r}
#adding a column for Julian date (might not be important, we'll see)
weekly_and_tony$julian_date <- julian(weekly_and_tony$date) 
#other method for julian date
#as.numeric(format(weekly_and_tony$date, "%j"))
```



To perform an interrupted time series analysis, you need two additional columns in the dataframe: One containing a binary variable that is 0 if the date is pre-nominations announcement and 1 if it is after, and another containing a binary variable that is 0 if the date is pre-ceremony date and 1 if it is after. 


```{r}
#TO DO (maybe): get exact dates of noms announcement and ceremony for every year instead of defaulting to May 1 and June 5

#adding this column for noms annoucement
weekly_and_tony$post_noms <- with(weekly_and_tony,
                                  ifelse(weekly_and_tony$date < as.Date(
                                    paste(weekly_and_tony$year, 05, 01, sep = "-")), 0, 1))

#adding this column for ceremony date
weekly_and_tony$post_awards <- with(weekly_and_tony,
                                  ifelse(weekly_and_tony$date < as.Date(
                                    paste(weekly_and_tony$year, 06, 05, sep = "-")), 0, 1))
```


```{r}
#add column for number of weeks a show has been open on any given week
#must order rows by date first 
weekly_and_tony <- weekly_and_tony[order(weekly_and_tony$date), ]

#add column 
weekly_and_tony$index <- ave(weekly_and_tony$production, FUN = seq_along) 
weekly_and_tony$weeks_open <- as.numeric(ave(weekly_and_tony$index, weekly_and_tony$production, FUN = seq_along))
#remove unnecessary index column
weekly_and_tony = subset(weekly_and_tony, select = -c(index))
```


Now I need to change the values of the "nominations" and "wins" columns to reflect the fact that prior to the date of the announcement, all productions had 0 nominations, and prior to the date of the ceremony, productions had 0 wins.

```{r}
#implementing this change
weekly_and_tony$nominations[weekly_and_tony$post_noms == 0] <- 0
weekly_and_tony$wins[weekly_and_tony$post_awards == 0] <- 0
```


```{r}
#want to analyze each year separately first
split_data_unclean <- split(weekly_and_tony, f = weekly_and_tony$year)
split_data <- list()

#splitting data changes column names for each year of data. To avoid having to deal with this later, we should change the column names back to what they originally were
cols_list <- colnames(weekly_and_tony)
for ( i in 1: length(split_data_unclean)){
  #change class to data frame for each year
  df <- as.data.frame(split_data_unclean[i])
  names(df) <- c(cols_list)
  #appending df to list where all dfs have the same column names
  split_data <- append(split_data, list(df))
}

#split_data
```


### segmented regression

Note that after writing the code for the segmented regression, I decided it was not a viable form of analysis for this project. I am keeping the code for now because I think it is worth preserving my thought process. The kind of segmented regression I originally wrote code for finds breakpoints on its own. However, for this project, we know the breakpoints a priori; They are the date of the nominations announcement and the date of the ceremony. Therefore, we want a model that has pre-specified breakpoints. 


```{r}
#testing this on a random year (2019)
test <- as.data.frame(split_data[40])
#names(test) <- c(cols_list)


#this graph seems too hectic to include in final report. Probably better to stick to more anecdotal things that you made in previous part, at least for visuals
ggplot(test, aes(x = date, y = capacity, color = production)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%b",
                 limits = as.Date(c('2018-05-01','2019-09-01'))) +
  #ggtitle("") +
  theme_bw() +
  theme(legend.position = "none")
```



```{r}
aint_proud <- test[test$production == 'aint-too-proud-520004',] 
test_lm <- lm(capacity ~ julian_date + nominations + wins + weeks_open, data = aint_proud)
segmented_test_lm <- segmented(test_lm, seg.Z = ~julian_date, 
                               psi = julian(as.Date("2019-05-01")))

#just for one random musical (Ain't too Proud)
summary(segmented_test_lm)
```



```{r}
#these results for segmented regression seem more accurate than they were before (i.e. when I did not include weeks_open. this var is very relevant to capacity)
test_lm <- lm(capacity ~ julian_date + nominations + wins + weeks_open, data = test)
segmented_test_lm <- segmented(test_lm, seg.Z = ~julian_date, 
                               psi = c(julian(as.Date("2019-05-01")),
                                       julian(as.Date("2019-06-05"))))

#estimates breakpoint as being right after awards (June 15, 2019)
summary(segmented_test_lm)
```




```{r}
#extracting two breakpoint dates from the model 
breaks_julian <- segmented_test_lm$indexU$date
#how to convert Julian date back into regular date (the default origin is Jan 1, 1970)
breaks <- as.Date(breaks_julian, origin = "1970-01-01")
#coefficients for all vars
coeffs <- segmented_test_lm$coefficients
#p-values for all vars
p_vals <- summary(segmented_test_lm)$coefficients[,4] 
#adjusted r-squared
r_sqr <- summary(segmented_test_lm)$adj.r.squared

#breaks_julian
#breaks
#coeffs
#coeffs["nominations"]
#p_vals
#r_sqr
```


```{r}
breaks_lol <- list()
coeffs_lol <- list()
signif_lol <- list()

for (df in split_data){
   # data for 1980 starts on date of 1980 Tonys (June 8, 1980), so there will not be breakpoints for this first model
  if (df$year[1] == 1980){
    segmented_year_lm <- lm(capacity ~ date + nominations + wins + weeks_open, data = df)
  }
  else{
    year_lm <- lm(capacity ~ date + nominations + wins + weeks_open, data = df)
    segmented_year_lm <- segmented(year_lm, seg.Z = ~date, 
                                  psi = c(as.Date(paste(df$year[1], 05, 01, sep = "-")),
                                          as.Date(paste(df$year[1], 06, 05, sep = "-"))))
  }

  breaks_julian <- segmented_year_lm$indexU$date
  breaks <- as.Date(breaks_julian, origin = "1970-01-01")
  coeffs <- segmented_year_lm$coefficients
  p_vals <- summary(segmented_year_lm)$coefficients[,4] 
  r_sqr <- summary(segmented_year_lm)$adj.r.squared
  
  breaks_list <- c(breaks)
  breaks_lol <- append(breaks_lol, list(breaks_list))
  
  year_coeffs <- c(coeffs, df$year[1])
  coeffs_lol <- append(coeffs_lol, list(year_coeffs))
  
  year_signif <- c(p_vals, r_sqr, df$year[1])
  signif_lol <- append(signif_lol, list(year_signif))
  
  #split_data <- append(split_data, list(df)) 
}
#as.Date(paste(year, 05, 01, sep = "-"))
```



```{r}
#summary(segmented_year_lm)
```

### Interrupted time series

Before I do the next part of my analysis, I want to do a simple linear regression to examine the relationship between the number of weeks since a production opened and the productions capacity in that week.


```{r}
#removing outliers barely affected results

ggplot(weekly_and_tony, aes(x = weeks_open, y = capacity)) +
  geom_jitter(color = 'maroon') +
  #not sure why line of best fit won't show up
  geom_smooth(method = "lm", color = 'skyblue') +
  labs(x = "weeks since opening", y = "weekly capcity") +
  ggtitle("Effect of Weeks Since Opening on Weekly Capacity") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "None")
```


```{r}
capacity_lm <- lm(capacity ~ weeks_open, data = weekly_and_tony)
summary(capacity_lm)
CI <- confint(capacity_lm)
CI
```

While the effect of weeks since opening on capacity is statistically significant, it is actually a significantly smaller effect than expected, and it is almost negligible. It may not be worth including in my time series model


```{r}
#mmr means moderated multiple regression
mmr <- lm(capacity ~ nominations*weeks_open + wins*weeks_open, data = weekly_and_tony)

summary(mmr)
```


```{r}
confint(mmr)
```

```{r}
plot(mmr)
```


```{r}
#does not meet homoscedasticity assumptions
ggplot(mmr, aes(x = mmr$fitted.values, y = mmr$residuals)) +
  geom_point(color = "maroon") +
  geom_hline(yintercept = 0, color = "skyblue") +
  labs(x = "fitted values", y = "residuals") +
  ggtitle("Residuals vs Fitted\n(for moderated multiple regression)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) 
```


```{r}
QQ_plot <- ggplot(mmr, aes(sample = mmr$residuals)) +
  stat_qq(color = "maroon") +
  stat_qq_line(color = "skyblue") + 
  labs(x = "theoretical quantiles", y = "sample quantiles") +
  ggtitle("Normal Q-Q Plot\n(for moderated multiple regression)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

resid_hist <- ggplot(mmr, aes(x = mmr$resid)) +
  geom_histogram(color = "black", fill = "skyblue") +
  #xlim(-500,500) +
  labs(x = "residual value", y = "frequency") +
  ggtitle("Distribution of Residuals\n(for moderated multiple regression)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(QQ_plot, resid_hist, ncol = 1, nrow = 2, labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/QQ+hist_mmr.png", bg = "transparent", height = 10, width = 8)
```


```{r}
#Interrupted time series
#testing on just one production first
#aint_proud_tm <- lm(capacity ~ julian_date + post_noms + post_awards + weeks_open + nominations + wins, data = aint_proud)

aint_proud_tm <- lm(capacity ~ nominations + wins + weeks_open, data = aint_proud)

#results indicate that for this show, the weeks it been open negatively affect its capacity (people want to see new things), time after noms led to a 4.8% increase in capacity, time after awards led to a 4.0% increase in capacity (not entirely accurate if you look at actual data for this show. Capacity did increase, but not by that much)
#did not deem number of Tony noms or wins relevant, probably because I am only looking at one show, so number of noms/wins almost never changes
summary(aint_proud_tm)
```


```{r}
#do the same thing as the above chunk but for every 2018-2019 production
#make post_noms/awards into one column 
#test_tm <- lm(capacity ~ julian_date + post_noms + post_awards + weeks_open + nominations + wins, data = test)

#ADD VAR that is weeks post awards???
test_tm <- lm(capacity ~ nominations*weeks_open + wins*weeks_open, data = test)

summary(test_tm)
```


```{r}
ts_coeffs <- list()
ts_signif <- list()
for (df in split_data){
  #longevity_lm <- lm(weeks ~ nominations + wins, data = df)
  #ts <- lm(capacity ~ julian_date + post_noms + post_awards + 
  #          weeks_open + nominations + wins, data = df)
  
  ts <- lm(capacity ~ nominations*weeks_open + wins*weeks_open, data = df)
  #ts <- lm(capacity ~ nominations + wins + weeks_open, data = df)
  
  coeffs <- c(ts$coefficients, df$year[1])
  ts_coeffs <- append(ts_coeffs, list(coeffs))
  p_vals <- summary(ts)$coefficients[,4] 
  r_sqr <- summary(ts)$adj.r.squared
  CI <- confint(ts, level=0.95)
  signif_list <- c(p_vals, r_sqr, CI, df$year[1])
  ts_signif <- append(ts_signif, list(signif_list))
}
```


```{r}
#making the results of the above chunk into a dataframe and making sure the column names make sense
ts_coeffs_df <- as.data.frame(do.call(rbind, ts_coeffs))
colnames(ts_coeffs_df) <- c("intercept_coeff", "noms_coeff", "weeks_open_coeff",
                              "wins_coeff", "noms_weeks_coeff", "wins_weeks_coeff", "year")

ts_signif_df <- as.data.frame(do.call(rbind, ts_signif))
colnames(ts_signif_df) <- c("intercept_pval", "noms_pval",  "weeks_open_pval", 
                            "wins_pval", "noms_weeks_pval", "wins_weeks_pval",
                            "rsqr", "CI_1_int", "CI_1_noms", "CI_1_weeks", "CI_1_wins",
                            "CI_1_noms_weeks", "CI_1_wins_weeks", 
                            "CI_2_int", "CI_2_noms", "CI_2_weeks", "CI_2_wins",
                            "CI_2_noms_weeks", "CI_2_wins_weeks", "year")

ts_df <- merge(ts_coeffs_df, ts_signif_df)
```


```{r}
#plotting all of the results
wins_effect_capacity <- ggplot(ts_df, aes(x = wins_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_wins, xmax = CI_2_wins), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "wins effect size") +
  ggtitle("Effect Size of Wins on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

noms_effect_capacity <- ggplot(ts_df, aes(x = noms_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_noms, xmax = CI_2_noms), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "nominations effect size") +
  ggtitle("Effect Size of Nominations on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

weeks_effect_capacity <- ggplot(ts_df, aes(x = weeks_open_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_weeks, xmax = CI_2_weeks), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "weeks since opening effect size") +
  ggtitle("Effect Size of Weeks Since Opening on\nCapacity for Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

noms_weeks_interaction <- ggplot(ts_df, aes(x = noms_weeks_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_noms_weeks, xmax = CI_2_noms_weeks), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "effect size of the interaction between Nominations and weeks since opening") +
  ggtitle("Effect Size of the Interaction Between Weeks\nSince Opening and Nominations on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

wins_weeks_interaction <- ggplot(ts_df, aes(x = wins_weeks_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_wins_weeks, xmax = CI_2_wins_weeks), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "effect size of the interaction between wins and weeks since opening") +
  ggtitle("Effect Size of the Interaction Between Weeks\nSince Opening and Wins on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(noms_effect_capacity, wins_effect_capacity, weeks_effect_capacity, align = "none", labels = "AUTO")
#save image to folder in repo
#ggsave("visuals/noms_wins_effect_capacity.png", bg = "transparent", height = 10, width = 10)

ggarrange(noms_weeks_interaction, wins_weeks_interaction, ncol = 1, nrow = 2, align = "none", labels = "AUTO")
#save image to folder in repo
#ggsave("visuals/interaction_effect_capacity.png", bg = "transparent", height = 10, width = 8)
```


```{r}
ggplot(ts_df, aes(x = weeks_open_coeff)) +
  geom_histogram(color = "black", fill = "skyblue") +
  geom_vline(xintercept = 0,lwd=.5,colour="maroon", linetype = 2) +
  ggtitle("Distribution of Yearly Regression Model Slopes for Weeks Open") +
  labs(x = "weeks open effect size") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```



```{r}
#relevant to note that adjusted r-squared indicates a poor fit
mean(ts_df$rsqr)
median(ts_df$rsqr)
```


