---
title: "Main Tony Analysis"
author: "Maximilian Wisnefski"
date: '2023-10-30'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(segmented)
library(ggpubr)
#library(nlme) for nonlinear mixed effects, could be useful in the future?
```


```{r}
#filtered data
weekly_and_tony <- read_csv("data/combined_filtered.csv", show_col_types = FALSE)
production_longevity <- read_csv("data/production_longevity.csv", show_col_types = FALSE)
```


## Longevity Analysis

This is a continuation of the regression analysis done in initial_Tony_analysis.Rmd. However, instead of creating one model, we will make a separate model for each year of data. 

```{r}
#getting rid of specials since their longevity is fixed and cannot be impacted by Tony outcomes
production_longevity <- production_longevity[production_longevity$type != "special", ]
#split data to analyze each year separately
split_data_unclean <- split(production_longevity, f = production_longevity$year)
split_longev <- list()

#splitting data changes column names for each year of data. To avoid having to deal with this later, we should change the column names back to what they originally were
cols_list <- colnames(production_longevity)
for ( i in 1: length(split_data_unclean)){
  #change class to dataframe for each year
  df <- as.data.frame(split_data_unclean[i])
  names(df) <- c(cols_list)
  #appending df to list where all dfs have the same column names
  split_longev <- append(split_longev, list(df))
}
```


```{r}
#creating lists to store the results of all 42 models
#this list stores coefficients (effect sizes) for predictors and intercept
long_coeffs <- list()

#this one stores all p-values, r-squared, and confidence intervals
long_signif <- list()

for (df in split_longev){
  #running the model for each year
  longevity_lm <- lm(weeks ~ nominations + wins, data = df)
  coeffs <- c(longevity_lm$coefficients, df$year[1])
  #storing output in long_coeffs
  long_coeffs <- append(long_coeffs, list(coeffs))
  p_vals <- summary(longevity_lm)$coefficients[,4] 
  r_sqr <- summary(longevity_lm)$adj.r.squared
  CI <- confint(longevity_lm, level=0.95)
  signif_list <- c(p_vals, r_sqr, CI, df$year[1])
  #storing output in long_signif
  long_signif <- append(long_signif, list(signif_list))
}
```


```{r}
#making the lists into dataframes and changing column names to things that make sense
long_coeffs_df <- as.data.frame(do.call(rbind, long_coeffs))
colnames(long_coeffs_df) <- c("intercept_coeff", "noms_coeff", "wins_coeff", "year")

long_signif_df <- as.data.frame(do.call(rbind, long_signif))
colnames(long_signif_df) <- c("intercept_pval", "noms_pval", "wins_pval", 
                              "rsqr", "CI_1_int", "CI_1_noms", "CI_1_wins", 
                              "CI_2_int", "CI_2_noms", "CI_2_wins", "year")

#merging the two dataframes
long_lm_df <- merge(long_coeffs_df, long_signif_df)
```


The next chunk creates visuals that allow us to see if and to what extent the impact of nominations/wins (on longevity) has changed over the last 40 years.

```{r}
#visualizing the results of the 42 models
wins_effect_plot <- ggplot(long_lm_df, aes(x = wins_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_wins, xmax = CI_2_wins), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "wins effect size") +
  ggtitle("Effect Size of Wins on Longevity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

noms_effect_plot <- ggplot(long_lm_df, aes(x = noms_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_noms, xmax = CI_2_noms), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "nominations effect size") +
  ggtitle("Effect Size of Nominations on Longevity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(noms_effect_plot, wins_effect_plot, ncol = 1, nrow = 2,  labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/noms_wins_effect_longev.png", bg = "transparent", height = 10, width = 8)
```

It is important to note that the years that produced models with very large confidence intervals are all years that had an outlier production that stayed open for an unusually long time. 1988 has Phantom of the Opera, 1998 has The Lion King, 2004 has Wicked, and 2014 has Aladdin. 


The histograms below show us the distribution of effect sizes for nominations/wins across all of the models.

```{r}
#visualizing the results of the 42 models
noms_effect_hist_longev <- ggplot(long_lm_df, aes(x = noms_coeff)) +
  geom_histogram(color = "black", fill = "skyblue") +
  geom_vline(xintercept = 0,lwd=.5,colour="maroon", linetype = 2) +
  ggtitle("Distribution of Yearly Regression Model Slopes for Nominations") +
  labs(x = "nominations effect size") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

wins_effect_hist_longev <- ggplot(long_lm_df, aes(x = wins_coeff)) +
  geom_histogram(color = "black", fill = "skyblue") +
  geom_vline(xintercept = 0,lwd=.5,colour="maroon", linetype = 2) +
  ggtitle("Distribution of Yearly Regression Model Slopes for Wins") +
  labs(x = "wins effect size") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(noms_effect_hist_longev, wins_effect_hist_longev, ncol = 1, nrow = 2,  labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/noms_wins_effect_hists.png", bg = "transparent", height = 10, width = 8)
```


```{r}
#many of the models only fit the data moderately well
mean(long_lm_df$rsqr)
median(long_lm_df$rsqr)
```


## Formatting Data for Segmented Regression and Moderated Multiple Regression

NOTE: Segmented regression did not end up being part of this analysis because it assumes you do not know the breakpoints a priori. However, we do know the breakpoints a priori; they are the date of nomination announcements, and the date of the award ceremony. I have kept the code for segmented regression for now because it seems wise to not delete the code even if I am not currently planning on using it.

```{r}
#adding a column for Julian date (did not end up being necessary, but we will keep it anyway)
weekly_and_tony$julian_date <- julian(weekly_and_tony$date) 
#other method for julian date
#as.numeric(format(weekly_and_tony$date, "%j"))
```


The next chunk was intended to be relevant to the interrupted time series, but I ended up performing moderated multiple regression instead. We will keep this chunk for now, as it could be useful in the next month. The chunk adds two new columns to the data: a column that is 0 if the date is pre-nomination announcement and 1 if the date is post-announcement, and a column that is 0 if the date is pre-ceremony, and 1 if the date is post-ceremony. 


```{r}
#adding this column for noms annoucement
weekly_and_tony$post_noms <- with(weekly_and_tony,
                                  ifelse(weekly_and_tony$date < as.Date(
                                    paste(weekly_and_tony$year, 05, 01, sep = "-")), 0, 1))

#adding this column for ceremony date
weekly_and_tony$post_awards <- with(weekly_and_tony,
                                  ifelse(weekly_and_tony$date < as.Date(
                                    paste(weekly_and_tony$year, 06, 05, sep = "-")), 0, 1))
```


The next chunk adds a column that specifies the number of weeks that have passed since a production opened for each week of data. This will be important when we do moderated multiple regression

```{r}
#add column for number of weeks a show has been open on any given week
#must order rows by date first 
weekly_and_tony <- weekly_and_tony[order(weekly_and_tony$date), ]

#add column 
weekly_and_tony$index <- ave(weekly_and_tony$production, FUN = seq_along) 
weekly_and_tony$weeks_open <- as.numeric(ave(weekly_and_tony$index, weekly_and_tony$production, FUN = seq_along))
#remove unnecessary index column
weekly_and_tony = subset(weekly_and_tony, select = -c(index))
```


Now we need to change the values of the "nominations" and "wins" columns to reflect the fact that prior to the date of the announcement, all productions for a given year had 0 nominations, and prior to the date of the ceremony, all productions for a given year had 0 wins.

```{r}
#implementing this change
weekly_and_tony$nominations[weekly_and_tony$post_noms == 0] <- 0
weekly_and_tony$wins[weekly_and_tony$post_awards == 0] <- 0
```


```{r}
#want to analyze each year separately first
split_data_unclean <- split(weekly_and_tony, f = weekly_and_tony$year)
split_data <- list()

#splitting data changes column names for each year of data. To avoid having to deal with this later, we should change the column names back to what they originally were
cols_list <- colnames(weekly_and_tony)
for ( i in 1: length(split_data_unclean)){
  #change class to data frame for each year
  df <- as.data.frame(split_data_unclean[i])
  names(df) <- c(cols_list)
  #appending df to list where all dfs have the same column names
  split_data <- append(split_data, list(df))
}

#split_data
```


## Segmented Regression

NOTE (again): after writing the code for the segmented regression, I decided it was not a viable form of analysis for this project. I am keeping the code for now because I think it is worth preserving my thought process, and some of the code could be useful in the next month. The kind of segmented regression I originally wrote code for finds breakpoints on its own. However, for this project, we know the breakpoints a priori.


```{r}
#testing this on a random year (2019)
test <- as.data.frame(split_data[40])
#names(test) <- c(cols_list)

#plotting weekly capacity for all shows from the 2018 - 2019 season
ggplot(test, aes(x = date, y = capacity, color = production)) +
  geom_line() +
  scale_x_date(date_breaks = "1 month", 
                 date_labels = "%b",
                 limits = as.Date(c('2018-05-01','2019-09-01'))) +
  #ggtitle("") +
  theme_bw() +
  theme(legend.position = "none")
```

Note that the above visual is not in the final paper. It is too difficult to interept what's going on here and too difficult to distinguish between different productions


```{r}
#making dataframe for just one musical
aint_proud <- test[test$production == 'aint-too-proud-520004',] 
test_lm <- lm(capacity ~ julian_date + nominations + wins + weeks_open, data = aint_proud)
segmented_test_lm <- segmented(test_lm, seg.Z = ~julian_date, 
                               psi = julian(as.Date("2019-05-01")))

#just for one random musical (Ain't too Proud)
summary(segmented_test_lm)
```



```{r}
#these results for segmented regression seem more accurate than they were before (i.e. when I did not include weeks_open. this variable seems relevant to capacity)
test_lm <- lm(capacity ~ julian_date + nominations + wins + weeks_open, data = test)
segmented_test_lm <- segmented(test_lm, seg.Z = ~julian_date, 
                               psi = c(julian(as.Date("2019-05-01")),
                                       julian(as.Date("2019-06-05"))))

#estimates breakpoint as being right after awards (June 15, 2019)
summary(segmented_test_lm)
```


```{r}
#extracting two breakpoint dates from the model 
breaks_julian <- segmented_test_lm$indexU$date
#how to convert Julian date back into regular date (the default origin is Jan 1, 1970)
breaks <- as.Date(breaks_julian, origin = "1970-01-01")
#coefficients for all vars
coeffs <- segmented_test_lm$coefficients
#p-values for all vars
p_vals <- summary(segmented_test_lm)$coefficients[,4] 
#adjusted r-squared
r_sqr <- summary(segmented_test_lm)$adj.r.squared

#breaks_julian
#breaks
#coeffs
#coeffs["nominations"]
#p_vals
#r_sqr
```


In the next chunk, I began the process of gathering the results of all 42 segmented regression models into one dataframe, but before I finished, I decided not to use this form of analysis. Therefore, this next chunk is a bit unfinished.

```{r}
breaks_lol <- list()
coeffs_lol <- list()
signif_lol <- list()

for (df in split_data){
   # data for 1980 starts on date of 1980 Tonys (June 8, 1980), so there will not be breakpoints for this first model
  if (df$year[1] == 1980){
    segmented_year_lm <- lm(capacity ~ date + nominations + wins + weeks_open, data = df)
  }
  else{
    year_lm <- lm(capacity ~ date + nominations + wins + weeks_open, data = df)
    segmented_year_lm <- segmented(year_lm, seg.Z = ~date, 
                                  psi = c(as.Date(paste(df$year[1], 05, 01, sep = "-")),
                                          as.Date(paste(df$year[1], 06, 05, sep = "-"))))
  }

  breaks_julian <- segmented_year_lm$indexU$date
  breaks <- as.Date(breaks_julian, origin = "1970-01-01")
  coeffs <- segmented_year_lm$coefficients
  p_vals <- summary(segmented_year_lm)$coefficients[,4] 
  r_sqr <- summary(segmented_year_lm)$adj.r.squared
  
  breaks_list <- c(breaks)
  breaks_lol <- append(breaks_lol, list(breaks_list))
  
  year_coeffs <- c(coeffs, df$year[1])
  coeffs_lol <- append(coeffs_lol, list(year_coeffs))
  
  year_signif <- c(p_vals, r_sqr, df$year[1])
  signif_lol <- append(signif_lol, list(year_signif))
  
  #split_data <- append(split_data, list(df)) 
}
#as.Date(paste(year, 05, 01, sep = "-"))
```



```{r}
#summary(segmented_year_lm)
```

## Moderated Multiple Regression


```{r}
#mmr stands for moderated multiple regression
mmr <- lm(capacity ~ nominations*weeks_open + wins*weeks_open, data = weekly_and_tony)

summary(mmr)
```


```{r}
#finding confidence intervals
confint(mmr)
```

```{r}
#checking assumptions
plot(mmr)
```


```{r}
#does not meet homoscedasticity assumptions
ggplot(mmr, aes(x = mmr$fitted.values, y = mmr$residuals)) +
  geom_point(color = "maroon") +
  geom_hline(yintercept = 0, color = "skyblue") +
  labs(x = "fitted values", y = "residuals") +
  ggtitle("Residuals vs Fitted\n(for moderated multiple regression)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5)) 
```


```{r}
#does not meet normality assumptions
QQ_plot <- ggplot(mmr, aes(sample = mmr$residuals)) +
  stat_qq(color = "maroon") +
  stat_qq_line(color = "skyblue") + 
  labs(x = "theoretical quantiles", y = "sample quantiles") +
  ggtitle("Normal Q-Q Plot\n(for moderated multiple regression)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

resid_hist <- ggplot(mmr, aes(x = mmr$resid)) +
  geom_histogram(color = "black", fill = "skyblue") +
  #xlim(-500,500) +
  labs(x = "residual value", y = "frequency") +
  ggtitle("Distribution of Residuals\n(for moderated multiple regression)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(QQ_plot, resid_hist, ncol = 1, nrow = 2, labels = "AUTO")

#save image to folder in repo
#ggsave("visuals/QQ+hist_mmr.png", bg = "transparent", height = 10, width = 8)
```


```{r}
#moderated multiple regression
#testing on just one production first
aint_proud_mmr <- lm(capacity ~ nominations + wins + weeks_open, data = aint_proud)

summary(aint_proud_mmr)
```


```{r}
#do the same thing as the above chunk but for every 2018-2019 production
test_mmr <- lm(capacity ~ nominations*weeks_open + wins*weeks_open, data = test)

summary(test_mmr)
```


```{r}
mmr_coeffs <- list()
mmr_signif <- list()
for (df in split_data){
  #running the model for each year of data
  mmr <- lm(capacity ~ nominations*weeks_open + wins*weeks_open, data = df)
  
  #storing output in mmr_coeffs
  coeffs <- c(mmr$coefficients, df$year[1])
  mmr_coeffs <- append(mmr_coeffs, list(coeffs))
  
  #storing output in mmr_signif
  p_vals <- summary(mmr)$coefficients[,4] 
  r_sqr <- summary(mmr)$adj.r.squared
  CI <- confint(mmr, level=0.95)
  signif_list <- c(p_vals, r_sqr, CI, df$year[1])
  mmr_signif <- append(mmr_signif, list(signif_list))
}
```


```{r}
#making the results of the above chunk into a dataframe and making sure the column names make sense
mmr_coeffs_df <- as.data.frame(do.call(rbind, mmr_coeffs))
colnames(mmr_coeffs_df) <- c("intercept_coeff", "noms_coeff", "weeks_open_coeff",
                              "wins_coeff", "noms_weeks_coeff", "wins_weeks_coeff", "year")

mmr_signif_df <- as.data.frame(do.call(rbind, mmr_signif))
colnames(mmr_signif_df) <- c("intercept_pval", "noms_pval",  "weeks_open_pval", 
                            "wins_pval", "noms_weeks_pval", "wins_weeks_pval",
                            "rsqr", "CI_1_int", "CI_1_noms", "CI_1_weeks", "CI_1_wins",
                            "CI_1_noms_weeks", "CI_1_wins_weeks", 
                            "CI_2_int", "CI_2_noms", "CI_2_weeks", "CI_2_wins",
                            "CI_2_noms_weeks", "CI_2_wins_weeks", "year")

mmr_df <- merge(mmr_coeffs_df, mmr_signif_df)
```


The next chunk creates visuals that allow us to see if and to what extent the impact of nominations/wins (on weekly capacity) has changed over the last 40 years.

```{r}
#plotting all of the results
wins_effect_capacity <- ggplot(mmr_df, aes(x = wins_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_wins, xmax = CI_2_wins), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "wins effect size") +
  ggtitle("Effect Size of Wins on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

noms_effect_capacity <- ggplot(mmr_df, aes(x = noms_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_noms, xmax = CI_2_noms), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "nominations effect size") +
  ggtitle("Effect Size of Nominations on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

weeks_effect_capacity <- ggplot(mmr_df, aes(x = weeks_open_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_weeks, xmax = CI_2_weeks), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "weeks since opening effect size") +
  ggtitle("Effect Size of Weeks Since Opening on\nCapacity for Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

noms_weeks_interaction <- ggplot(mmr_df, aes(x = noms_weeks_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_noms_weeks, xmax = CI_2_noms_weeks), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "effect size of the interaction between Nominations and weeks since opening") +
  ggtitle("Effect Size of the Interaction Between Weeks\nSince Opening and Nominations on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

wins_weeks_interaction <- ggplot(mmr_df, aes(x = wins_weeks_coeff, y = year)) +
  geom_vline(xintercept = 0,lwd=1,colour="black", linetype = 2) +
  geom_errorbar(aes(xmin = CI_1_wins_weeks, xmax = CI_2_wins_weeks), alpha = 2,color = "skyblue") + 
  geom_point(color = "maroon") +
  labs(x = "effect size of the interaction between wins and weeks since opening") +
  ggtitle("Effect Size of the Interaction Between Weeks\nSince Opening and Wins on Capacity\nfor Every Year (1980-2023)") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

ggarrange(noms_effect_capacity, wins_effect_capacity, weeks_effect_capacity, align = "none", labels = "AUTO")
#save image to folder in repo
#ggsave("visuals/noms_wins_effect_capacity.png", bg = "transparent", height = 10, width = 10)

ggarrange(noms_weeks_interaction, wins_weeks_interaction, ncol = 1, nrow = 2, align = "none", labels = "AUTO")
#save image to folder in repo
#ggsave("visuals/interaction_effect_capacity.png", bg = "transparent", height = 10, width = 8)
```


```{r}
#did not end up including this visual or others like it in final analysis because it felt like overkill to have 5 more histograms in addition to the 5 plots above. Keeping code for now in case I decide to include it later
ggplot(mmr_df, aes(x = weeks_open_coeff)) +
  geom_histogram(color = "black", fill = "skyblue") +
  geom_vline(xintercept = 0,lwd=.5,colour="maroon", linetype = 2) +
  ggtitle("Distribution of Yearly Regression Model Slopes for Weeks Open") +
  labs(x = "weeks open effect size") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
```


```{r}
#relevant to note that adjusted r-squared indicates a very poor fit
mean(mmr_df$rsqr)
median(mmr_df$rsqr)
```


